
#Your name, date and assignment 
---
title: "Assignment 1_AYK"
author: "Yao"
output: html_document
date: "2024-09-17"
---

#Load relevant software libraries
```{r}
install.packages("tidyverse")
library(tidyverse)
library(rio)
library(janitor)

#Software libraries loaded
```

#Load the data: https://github.com/wellsdata/CompText_Jour/blob/main/data/blackindex_master.csv
```{r}
Black_Index <- read.csv("https://raw.githubusercontent.com/wellsdata/CompText_Jour/main/data/blackindex_master.csv")

#Dataset imported using the read.csv command
```

#Using code, describe the number of rows and columns in the dataset
```{r}
dim(Black_Index)
nrow(Black_Index)
ncol(Black_Index)

#the dataset has a dimension of 1803 rows and 30 columns
```

#Create a table that displays a count of the top 15 newspaper_city entries
```{r}
Top_15 <- Black_Index %>%
count (newspaper_city) %>%
 arrange(desc(newspaper_city)) %>%
  slice_max(n, n=15)

#Table displaying newspaper_city entries
```

#Create a table that counts the different entries in the "newspaper_city" column
```{r}
Dif_entries <- Black_Index %>%
count (newspaper_city, sort = TRUE, name = "Count") %>%
  arrange(desc(Count))

#Table of counts of different entries in the "newspaper_city" column
```

#Create a table that counts all newspaper_city entries by year
```{r}
Entries_by_Yr <- Black_Index %>%
  count (newspaper_city, year, sort = TRUE, name = "Count") %>%
  arrange(desc(Count))

#Table showing newspaper_city entries count by year
```

#Create a simple column ggplot chart that shows the total entries by year
```{r}
library(dplyr)
library(ggplot2)
  Black_Index %>%
  group_by(year) %>%
count(year) %>%
  ggplot(aes(x=year, y=n, fill=n))+
  geom_col(position = "dodge")+
labs(title = "Graph of total entries by year",
       subtitle = "Data covering years captured",
       caption = "Graphic by Yao")
  
  #Graph showing total entries by year
```
  