---
title: "homework-basic-r-proficiency"
author: "Rob Wells"
date: "2024-08-23"
output: html_document
---

# Homework - Basic R Proficiency

This class exercise is a companion to the following chapter in Computational Text Analysis: [Ch. 11  Loading and Analyzing Data](https://wellsdata.github.io/CompText_Jour/docs/r-load-analyze-visualize.html)


#1: load tidyverse, rio and janitor libraries
```{r}
library(tidyverse)
library(rio)
library(janitor)

#Tidyverse, Rio and Janitor packages successfully loaded.
```


#2: Import Data using read.csv:

https://github.com/wellsdata/CompText_Jour/raw/main/data/mob_analysis_data.csv
```{r}
Mob<-read.csv("https://github.com/wellsdata/CompText_Jour/raw/main/data/mob_analysis_data.csv")

#Mob data imported through the csv file link provided.
```

#3: Provide dimensions: Gives number rows, then columns

```{r}
dim(Mob)
nrow(Mob)
ncol(Mob)

#The Data has 3836 rows and 7 columns.
```

#4: Show names of your columns

```{r}
colnames(Mob)

#the displayed column names are "X", "docname", "Newspaper", "Year", "Beginning.of.Passage", "keyword" and "End.of.Passage.
```

#5: Check data types

```{r}
str(Mob)

#Data type previewed using the string command. Data types are integers and characters.
```

#6: Display the first six rows

```{r}
head(Mob)

#First six rows displayed using the head command.
```
#7: Create a table counting the number of entries by year and display the top 5 results

```{r}
YrTab <- Mob %>%
  group_by(Year) %>%
  summarise(
    count_tracts = n()
  )
head(YrTab, n = 5)

#Table counting number of entries by year and displaying top 5 results.
```
#8: Create a table that counts the different entries in the "keyword" column

```{r}
KeywordTab <- Mob %>%
  group_by(keyword) %>%
  summarise(
    count_tracts = n()
  )
# Table for different entries in the "keyword" column.
```


#9: Create a simple column ggplot chart that shows the total entries by year

```{r}
ggplot(YrTab, aes(x = Year, y = count_tracts)) +
geom_col(fill = "steelblue") +
  labs(title = "Total Entries by Year", x = "Year", y = "Total Entries") +
  theme_minimal()
```

